---
title: "Pedigree Verification"
author: "Bryan"
date: "June 13, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

First lets install and load a couple of packages that we'll need. *pedigreemm* for calculating additive relationships and *proxy* for calculating genetic distances.

```{r}
install.packages("pedigreemm")
install.packages("proxy")
library(pedigreemm)
library(proxy)
```

To read a bit about each of these packages and learn about their functions:

```{r}
help(package="pedigreemm")
help(package="proxy")

```

## Load pedigrees from Potato Paper

Ok, time to work with real data. Download the csv files from the potato paper (https://link.springer.com/article/10.1007%2Fs12230-016-9556-y). They are included at the very end of the paper in the supplementary material section.

```{r}
#Potato pedigrees and Additive Relationships

# set working directory and read in pedigree csv file
setwd("/home/bryan/Downloads")
potato_ped <- read.csv("12230_2016_9556_MOESM1_ESM.csv", header = TRUE, check.names = FALSE)

# use editPed to order the pedigree by generation and make sure there is a row for each unique accession name, including parents
potatoP2= editPed(dam=potato_ped$Mother, sire=potato_ped$Father, label=potato_ped$Name)
head(potatoP2)

# read records corrected by editPed into a formal pedigree object
potatoP3= pedigree(potatoP2$sire, potatoP2$dam, potatoP2$label)
head(potatoP3)

# calculate matrix of additive relationship scores based on pedigree
potatoAmat <- getA(potatoP3)
head(potatoAmat)
```

## Load marker/dosage matrix from Potato Paper

```{r}
#Potato Markers and Genetic Distances

# read in marker/dosage csv file
potato_markers <- read.csv("12230_2016_9556_MOESM2_ESM.csv", header = TRUE, check.names = FALSE)

# transpose markers so accessions are rows and markers are columns
tpotato_markers <- t(potato_markers)

# column names that should form the header are now in first row. So replace the header with the first row
colnames(tpotato_markers) <- as.character(unlist(tpotato_markers[1,]))
tpotato_markers = tpotato_markers[-1, ]

# set class as numeric so all matrix values will work in calculations, including NAs
class(tpotato_markers) <- "numeric"

# calculate Euclidean distances
potato_marker_distances <- dist(tpotato_markers, method = "Euclidean")

# look at the structure of the result object - its not a matrix anymore, its a 'dist'! 
str(potato_marker_distances) 

# convert distances to a matrix so they are easier to manipulate
dist_matrix = as.matrix(potato_marker_distances)

# check dimensions of our matrices
dim(potatoAmat)
dim(dist_matrix)

# We have too many genotypes, so make a list of all the accessions in the pedigree matrix, and use it to subset the genetic distance matrix
include_list <- labels(potatoAmat[,"W8848-3"]) #
subset_matrix <- dist_matrix[rownames(dist_matrix) %in% include_list, colnames(dist_matrix) %in% include_list] 
dim(subset_matrix) # now it should match

```

## Recreate Potato Paper Graph with ggplot2

```{r}
# install and load ggplot2 (Grammar of Graphics package)
install.packages("ggplot2")
library(ggplot2)
help(ggplot)

# get values for W8848-3, excluding self comparisons, from matrices
x <- potatoAmat[!rownames(potatoAmat) %in% "W8848-3","W8848-3"]
y <- subset_matrix["W8848-3",!colnames(subset_matrix) %in% "W8848-3"]

# ensure both vectors are in same order by sorting with accession names
xord <- order(names(x))
x <- x[xord]
yord <- order(names(y))
y <- y[yord]

# create plot!
ggplot()+ggtitle("W8848–3 Plot recreated")+labs(x="Additive relationship",y="Marker-based distance")+geom_text(aes(x=x,y=y,label=names(x)))

```


## Now apply the same techniques to the data from Cassavabase!

```{r}

setwd("/home/bryan/Desktop") # set the working directory to where you mounted/copied/downloaded your files

ped <- read.table("Pedigrees.txt", sep="\t", header = TRUE)
markers <- read.table("genotypes.txt", header = TRUE)


```


